Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-16 foreground_delay_spectrum_modeling.tex
Judd's initial edits

Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission. With much of the information coming from the spectral line that maps to distance along line of sight (cosmic time), foregrounds affecting the spectral domain deserve close attention. We investigate in detail the chromatic signatures of an all--sky foreground model imprinted by its spectral structure and instrumental transfer function. We use an antenna pair based delay spectrum technique. Besides confirming foreground contamination is predominantly restricted to a {\it wedge}--shaped region in Fourier space, we demonstrate that the foreground signatures which have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array (MWA) observations with simulations broken down by sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor even on wide antenna spacings relative to compact objects, hitherto unpredicted. These roles are switched in and around the primary field of view. This imprints a {\it pitchfork} on the {\it wedge} in Fourier space. The observation that the brightest contaminants are localized and only effect a fraction of the baselines leads us to propose that by down--weighting certain baselines at certain times, a large fraction of foreground contamination can be mitigated.

@@

Sensitive instruments such as the Square Kilometer Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum.

Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}. Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}. A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here we will focus primarily on the {\it avoidance} strategy, focusing on identifying foreground components most likely to ``leak'' from foreground modes to EoR modes.

Foregrounds leak into and contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis. Here we use existing catalogs and a high fidelity instrumental model to focus on the possible sources of instrumental chromaticity. 
To decouple from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual antenna pair based approach of \citet{pob13,par14}. This approximates the power spectrum as the inverse fourier transform of the spectra generated by the correlator.

 Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline. We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon. We also find that the smooth galactic emission plays a larger role then previously thought. We quantify this contamination by breaking the simulations down by sky region and type of emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

@@

We now describe in detail an important ingredient, the MWA instrument.

\section{Murchison Widefield Array Observations}\label{sec:instrument}

\subsection{The MWA}

@@

One important feature of the MWA's measured spectrum is the shape of its bandpass. The band is divided coarsely into 24 chunks with each chunk weighted by a digital filter. The channels at the edges of these chunks have been flagged because they are known to possibly be contaminated by aliasing at a low level. Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}. The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.

@@

Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all antenna pairs throughout the course of observation. We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction. ({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?})

@@

The MWA Epoch of reionization targets two primary low-foreground fields. Here we choose to focus on RA = 0h, Dec = -30\arcdeg.  The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in a MWA tile. These beamformer delay settings can be changed only in discrete steps. The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky. This process is repeated throughout the course of the observation $\approx 4.86$~hours. Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.

The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013. This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of low number of known system errors from among a long stretch of stable observing.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
\caption{MWA tile beam pointing directions during the course of the observation. The x-axis refers to the Local Sidereal Time (LST) in hours. The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees. Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours. The axis on the right refers to the declination of the pointing direction. The RA, HA, and declination are plotted with solid, dashed and dotted line styles respectively. A ``drift and shift'' scheme is used. The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. Dot--dashed vertical lines show the LST of off-zenith and zenith pointings used in our study. \label{fig:pointings}}
\end{figure}

\subsection*{A note on baselines/delay spectra sign convention}

Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the MWA site. For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively. For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set. Signals from this direction are recorded with a negative delay on baselines whose antenna spacings are oriented towards the east. Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward oriented antenna spacings.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, in prep). MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning. It has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on a forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antenna to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay. Finally, we fit for an additional phase known to be caused by reflection in 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect. In delay--space, these weights translate into a convolution by a ringing harmonic point spread function (PSF). We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limits corresponding to smooth spectrum objects in the visible sky hemisphere.

@@

In subsequent plots, we plot the amplitude of the delay spectrum, $|V(|\overline{\mathbf{x}}|,\eta)|$, sorted by baseline length, $|\overline{\mathbf{x}}|$. Note that visibilities from different baselines have not been averaged together in this analysis.

We show results of delay spectrum from MWA data after deconvolution along delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases linearly with light travel time corresponding to the baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we are investigating.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis). For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly. This is predominant at short antenna spacings where the number of delay modes within the horizon is lesser. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool. Hence, our discussion and conclusions are valid despite imperfections in deconvolution.

\section{Instrument Modeling}\label{sec:modeling}

@@

The precise power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active investigation. The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power--pattern of the dipole modeled via finite element electromagnetic simulation. To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the FEKO (or just aforementioned?) simulation. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz. An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane. Random fluctutations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:power_pattern}}
\end{figure}

@@

The degree to which these objects leak power into spectral line of sight modes above the wedge, a region known as the ``EoR Window'', depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.

\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it EoR window}. \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic compact objects in the {\it EoR window} besides thermal noise. In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum. We use an all--sky foreground emission model that includes both diffuse and bright compact components.

@@

For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. Using the same model, we use the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of diffuse emission.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

@@

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their complimentary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
\caption{Flux densities (in Jy) of bright compact objects at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5. The locations of bright compact objects appear to be random.\label{fig:CSM}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

>>>>

Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.  Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor, even on wide antenna spacings.   For signals entering through the primary field of view, compact objects dominate the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the {\it wedge} in Fourier space.  Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate a large fraction of foreground contamination in redshifted 21~cm power spectrum analysis.

@@

Sensitive instruments such as the Square Kilometer Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13} {\bf tbd: and bowman et al. 2013~}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum. {\bf  [JDB: add citations to Bowman et al. 2006, Beardsley et al. 2013, and Pober?]}

Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}. Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} {\bf [JDB: add citation to Liu et al. 2009, 2010]}. A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.   To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline-based approach of \citet{pob13,par14}. This approximates the power spectrum as the inverse fourier transform of the spectra generated by the instrument's correlator.

Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We confirm {\bf [JDB: is confirm the right word, seems too weak}.  Who else has published this?  Are we the first to declare it?] that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather from near the horizon.  We find that diffuse Galactic emission plays a larger role then previously thought. We quantify this contamination by separating the simulations into componetes base on sky region and type of foreground emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site. For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively. For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set. Signals from this direction are recorded with a negative delay. Similarly, signals from directions eastward of the local meridian arrive with positive delays.

@@

\section{Example Data}\label{sec:instrument}

\subsection{The Murchison Widefield Array (MWA)}

@@

The MWA bandpass is divided coarsely into 24 chunks with each chunk weighted by a digital filter. The channels at the edges of these chunks have been flagged because they are known to possibly be contaminated by aliasing at a low level. Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}. The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.

@@

Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all antenna pairs throughout the course of observation. We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction. ({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction? -- JDB: yes})

@@

The MWA  targets two primary low-foreground fields for reionization observations. Here, we focus on the field at RA = 0h, Dec = -30\arcdeg.  The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in an MWA tile. These beamformer delay settings can be changed only in discrete steps. The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky. This process is repeated throughout the course of the observation $\approx 4.86$~hours. Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.

The observations shown here were observed on Aug 23, 2013. This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of low number of known system errors from among a long stretch of stable observing.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings}
\caption{MWA tile beam pointing directions during the course of the observation. The x-axis refers to the Local Sidereal Time (LST) in hours. The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees. Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours. The axis on the right refers to the declination of the pointing direction. The RA, HA, and declination are plotted with solid, dashed and dotted line styles respectively. A ``drift and shift'' scheme is used. The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. Dot--dashed vertical lines show the LST of off-zenith and zenith pointings used in our study. \label{fig:pointings}}
\end{figure}

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, in prep). MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning. It has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on a forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay. Finally, we fit for an additional phase known to be caused by reflection in 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function. The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect. In delay--space, these weights translate into a convolution by a ringing harmonic point spread function (PSF). We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay, corresponding to smooth spectrum objects in the visible sky hemisphere.

@@

In subsequent sections, we plot the amplitude of the delay spectrum, $|V(|\overline{\mathbf{x}}|,\eta)|$, sorted by baseline length, $|\overline{\mathbf{x}}|$. Note that visibilities from different baselines have not been averaged together in this analysis.

We show results of delay spectrum from MWA data after deconvolution along delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases linearly with light travel time corresponding to the baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we will explore below.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots}
\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis). For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly. This is predominant at short antenna spacings where the number of delay modes within the horizon is lesser. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool.

\section{Simulations}\label{sec:modeling}

@@

The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled {\bf [TBD: cite Sutinjo et al (2014; submitted?)]} as a mutually-coupled 4-by-4 dipole array with the overall power--pattern of each inidivual dipole calculated via finite element electromagnetic simulation. To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain {\bf [TBD: cite: bowman et al. 2007 -- mwa prototype paper]}.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom}
\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz. An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane. Random fluctutations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:power_pattern}}
\end{figure}

@@

The degree to which these objects leak power into spectral line of sight modes above the wedge, depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.

% JDB: I don't think we need this here. %% \citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it EoR window}. \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic compact objects in the {\it EoR window} besides thermal noise. In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum.

We use an all--sky foreground emission model that includes both diffuse and bright compact components.

@@

For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. We model the spectral dependence of the diffuse foreground as a single spectral index at each pixel in our map, found by comparing model maps at 170~MHz and 200~MHz.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dsm}
\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

@@

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact source catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/csm}
\caption{Flux densities (in Jy) of bright compact objects at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5. The locations of bright compact objects appear to be random.\label{fig:CSM}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties:
